---
title: Is there a general relationship between the seagrass _Zostera noltii_ biomass and the field
  spectroradiometer readings?
author: "Lee, Chengfa Benjamin"
date: "18 December 2025 (last updated on `r format(Sys.Date(), '%d %B %Y')`)"
output: bookdown::html_document2 # html_document (for publishing), bookdown::html_document2 (for publishing with dynamic referencing), html_notebook (for previewing)
bibliography: references.bib 
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(readxl)
library(signal) 
library(ggplot2)
library(ggiraph)
library(dplyr)
library(tidyr)
library(stringr)
library(ggpubr)
library(ggtext)
library(patchwork)
```

This study aims to establish a relationship between the amount of biomass and reflectance of the _Zostera noltii_ seagrass. 

## Method

This data exploration uses the hyperspectral and biomass data collected during these field campaigns in Aveiro ([July 2025](https://sigoiry.github.io/Spectra_Aveiro_2025/)), Bourgneuf Bay ([2018, 2019](https://coastobs.eu/success-story/french-atlantic-coast), and [September 2025](https://sigoiry.github.io/Fieldtrip_Rewrite_Sept2025/)), and Cadiz Bay ([June 2025](https://rewriteproject.eu/)). 

### Hyperspectral data

Full methods can be found [here](https://leechengfa.github.io/seagrass_reflectance_2025/).

```{r hsPreprocessingBB}
#| echo: false
#| error: false
#| message: false
#| warning: false

srf <- read_xlsx('data/COPE-GSEG-EOPG-TN-15-0007 - Sentinel-2 Spectral Response Functions 2022 - 3.2.xlsx', sheet = 'Spectral Responses (S2A)') %>% 
  filter(SR_WL >= 400, SR_WL <= 900)

csv.spectra.BB <- local({
  reflectance.2018 <- read.table('data/spectra/BB/2018_09_Bourgneuf_reflectance-cores_raw.txt', header = TRUE) %>%
    pivot_longer(cols = names(.)[-1], names_to = "id", values_to = 'raw.value') %>%
    rename(Wavelength = wavelength)
  
    reflectance.2018$id <- reflectance.2018$id %>%
      lapply(FUN = function(string) {
        string %>%
          strsplit(split = "\\.") %>% 
          unlist() %>%
          "["(1) %>%
          strsplit(split = "_") %>% 
          unlist() %>%
          "["(-length(.)) %>%
          paste(collapse = "_")
      }) %>%
      unlist()
    reflectance.2018$id <- paste(reflectance.2018$id, '2018', sep = ".")
    reflectance.2018$year <- 2018
    
  reflectance.2018 <- reflectance.2018 %>%
    group_by(id) %>%
    mutate(raw.value = sgolayfilt(raw.value, 3, 11), .keep = 'all') %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    ungroup()
  
  reflectance.2025 <- read.csv('data/spectra/BB/20250911_Bourgneuf_H_stations_reflectance.txt', sep = '\t') %>%
    pivot_longer(cols = names(.) %>% grep('Wavelength', ., invert = TRUE, value = TRUE), names_to = "ASD", values_to = 'value') 
    
    reflectance.2025$id <- lapply(reflectance.2025$ASD, FUN = function(string) {
      str_split(string, '\\.') %>%
        unlist() %>%
        "["(1)
      }) %>% 
      unlist()
  
  reflectance.2025 <- reflectance.2025 %>%
    select(-ASD) %>%
    group_by(Wavelength,id) %>%
    summarise(raw.value = mean(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = sgolayfilt(raw.value, 3, 11), .keep = 'all') %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    ungroup()
  
  reflectance.2025$id <- reflectance.2025$id %>%
    gsub('H', "H_", .)  
  reflectance.2025$year <- 2025
  
  reflectance.2018 %>% 
    rbind(reflectance.2025) 
})
```


```{r hsPreprocessingAV}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.spectra.AV <- local({
  df <- list.files('data/spectra/AV', full.names = TRUE) %>%
    grep('\\.txt', ., value = TRUE) %>%
    lapply(FUN = function(f) {
      g <- read.csv(f, sep = '\t')
      h <- names(g)[2] %>% 
        gsub('\\.asd', "", .)
        names(g)[2] <- "reading"
        g$Name <- h
      
      g
    }) %>%
    do.call(rbind, .)
  
  df$dataset <- df$Name %>%
    lapply(FUN = function(name) {
      out <- str_split(name, '00') %>% 
        unlist() %>%
        head(1) %>%
        unlist()
    }) %>% 
    unlist()
  df$rep <- df$Name %>%
    lapply(FUN = function(name) {
      out <- str_split(name, '00') %>% 
        unlist() %>%
        tail(1) %>%
        unlist()
    }) %>%
    unlist() %>%
    as.numeric() %>%
    "+"(1) 
  
  ## add in station IDs from metadata file
  df <- df %>%
    mutate(id = case_when(
    # NG group
    dataset == 'AVEI2' & rep >= 341 & rep <= 360 ~ "NG12",
    dataset == 'AVEI2' & rep >= 401 & rep <= 420 ~ "NG15",
    dataset == 'AVEI2' & rep >= 381 & rep <= 400 ~ "NG16",
    dataset == 'AVEI2' & rep >= 301 & rep <= 320 ~ "NG18",
    dataset == 'AVEI2' & rep >= 281 & rep <= 300 ~ "NG25",
    dataset == 'AVEI2' & rep >= 261 & rep <= 280 ~ "NG30",
    dataset == 'AVEI2' & rep >= 221 & rep <= 240 ~ "NG31",
    dataset == 'AVEI2' & rep >= 201 & rep <= 220 ~ "NG32",
    dataset == 'AVEI2' & rep >= 361 & rep <= 380 ~ "NR43", 
    dataset == 'AVEI2' & rep >= 321 & rep <= 340 ~ "NR44", 
    dataset == 'AVEI2' & rep >= 241 & rep <= 260 ~ "NR52", 

    # RR/RG group
    dataset == 'AVE' & rep >= 1 & rep <= 20 ~ "RG12",
    dataset == 'AVE' & rep >= 21 & rep <= 40 ~ "RR36",
    dataset == 'AVE' & rep >= 41 & rep <= 60 ~ "RG5", 
    dataset == 'A' & rep >= 1 & rep <= 20 ~ "RG13",
    dataset == 'A' & rep >= 21 & rep <= 40 ~ "RG6", 
    dataset == 'A' & rep >= 41 & rep <= 60 ~ "RG7", 
    dataset == 'A' & rep >= 61 & rep <= 80 ~ "RG8", 
    dataset == 'A' & rep >= 101 & rep <= 120 ~ "RR46",
    dataset == 'A' & rep >= 121 & rep <= 140 ~ "RG16",
    dataset == 'A' & rep >= 141 & rep <= 160 ~ "RG23",
    dataset == 'A' & rep >= 161 & rep <= 180 ~ "RR45",
    dataset == 'A' & rep >= 181 & rep <= 200 ~ "RG22"
  ))
  
  df <- df %>%
    # average the readings 
    filter(!is.na(id)) %>%
    group_by(id, Wavelength) %>%
    select(-c(rep, Name)) %>%
    summarise(raw.value = mean(reading), dataset = unique(dataset)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = sgolayfilt(raw.value, 3, 11), .keep = 'all') %>%
    filter(Wavelength >= 400 & Wavelength <= 900) %>%
    ungroup() 
  
  meta <- list.files("data/spectra/AV", full.names = TRUE) %>%
    grep('Metadata_spectra', ., value = TRUE) %>%
    read_xlsx() %>%
    select(-c(Spectra)) 
  
  df <- df %>%
    left_join(meta, by = join_by(id == ID))
})
```

```{r hsPreprocessingCB}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.spectra.CB <- local({
  reflectance <- read.csv('data/spectra/CB/20250610_rw_cadiz_seagrass.txt', sep = '\t') %>%
    pivot_longer(cols = names(.) %>% grep('Wavelength', ., invert = TRUE, value = TRUE), names_to = "ASD", values_to = 'value') %>%
    mutate(group = unlist(substr(ASD, 1, 4)), rep = as.numeric(unlist(substr(ASD, 7, 10))))
  
  readMe <- read.csv('data/spectra/CB/20250610_rw_cadiz_seagrass_readme.txt', sep = ',') %>%
    mutate(group = substr(Spectrum_id_start, 1, 4), 
           id_start = as.numeric(unlist(substr(Spectrum_id_start, 7, 10))), id_end = as.numeric(unlist(substr(Spectrum_id_end, 7, 10)))
           )
    
    reflectance$id <- reflectance %>% 
      apply(MARGIN = 1, FUN = function(row) {
        output <- readMe$Station[readMe$group == row['group'] & readMe$id_start <= as.numeric(row['rep']) & readMe$id_end >= as.numeric(row['rep'])]
        ifelse(is.null(output), NA, output)
      }) %>% 
      unlist()
  
  reflectance <- reflectance %>%
    select(-ASD) %>%
    filter(!is.na(id)) %>%
    group_by(Wavelength,id) %>%
    summarise(raw.value = mean(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = sgolayfilt(raw.value, 3, 11), .keep = 'all') %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    ungroup()
  
  reflectance$year <- 2025
  
  reflectance 
})
```

### Biomass data

Full methods can be found [here](https://leechengfa.github.io/seagrass_carbon_2025/).

```{r biomassPreprocessingBB}
#| echo: false
#| error: false
#| message: false
#| warning: false

convert_biomass_to_per_m2 <- function(biomass) {
  core_area <- pi*((15/2)^2)
  biomass_per_m2 <- biomass*(100^2/core_area)
  
  return(biomass_per_m2)
}

csv.biomass.BB <- local({
  biomass.2018 <- read.csv('data/biomass/BB/2018_09_Bourgneuf_cores_biomass_percent_cover.csv', sep = '\t') %>%
    rename(id = Sample_ID, A_per_m2 = `Leaf_dry_weight_g`) %>%
    select(-Percent_cover)
    biomass.2018$A_per_m2 <- convert_biomass_to_per_m2(biomass.2018$A_per_m2)
    biomass.2018$B_per_m2 <- NA
    biomass.2018$year <- 2018
    biomass.2018$id <- paste(biomass.2018$id, '2018', sep = ".")
    
  spectra.2018 <- csv.spectra.BB %>%
    filter(year == 2018) %>%
    group_by(id) %>%
    summarise(red = weighted.mean(raw.value, srf$S2A_SR_AV_B4), nir = weighted.mean(raw.value, srf$S2A_SR_AV_B8), ndvi = (nir - red)/(nir + red))  
  
  df.2018 <- biomass.2018 %>%
    left_join(spectra.2018, by = join_by(id == id)) %>%
    group_by(id)  %>%
    select(-red, -nir) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    rename(NDVI = ndvi)
  
  biomass.2019 <- read_xlsx('data/biomass/BB/Data AGB BGB Bourgneuf 2019.xlsx') %>%
    rename(id = Sample_ID, A_per_m2 = `Leaf biomass (g DW m-2)`, B_per_m2 = `Rizhome biomass  (g DW m-2)`
) %>%
    select(-`Seagrass coverage (%)`)
    biomass.2019$year <- 2019
    
  df.2019 <- biomass.2019 %>%
    group_by(id) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2')
    df.2019$id <- paste(df.2019$id, '2019', sep = ".")
  
  biomass.2025 <- read_xlsx('data/biomass/BB/DryWeight_Cores.xlsx') %>%
    rename(id = Station)
  
    biomass.2025$A_per_m2 <- convert_biomass_to_per_m2(biomass.2025$A)
    biomass.2025$B_per_m2 <- convert_biomass_to_per_m2(biomass.2025$B)
    biomass.2025$year <- 2025
    
  spectra.2025 <- csv.spectra.BB %>%
    filter(year == 2025) %>%
    group_by(id) %>%
    summarise(red = weighted.mean(raw.value, srf$S2A_SR_AV_B4), nir = weighted.mean(raw.value, srf$S2A_SR_AV_B8), ndvi = (nir - red)/(nir + red))  
  
  df.2025 <- biomass.2025 %>%
    left_join(spectra.2025, by = join_by(id == id)) %>%
    group_by(id)  %>%
    select(-red, -nir, -A, -B) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    rename(NDVI = ndvi)
  
  rbind(df.2018, df.2019, df.2025)
})
```

```{r biomassPreprocessingAV}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.AV <- local({
  df <- read_xlsx('data/biomass/AV/Aveiro_grid_info_final.xlsx', sheet = 'data') %>%
    select(-c(longitude, latitude, poskey.benthos, poskey.sediment, remark))
    df$`spectral camera`[df$`spectral camera` == 'no'] <- NA
    df$spectra <- !is.na(df$`spectral camera`)
    
    df$A_per_m2 <- convert_biomass_to_per_m2(df$seagrassA/1000)
    df$B_per_m2 <- convert_biomass_to_per_m2(df$seagrassB/1000)
    df$A_per_m2[df$B_per_m2 == 0] <- 0 # if there's no root biomass, the leaf biomass are not growing at this station
    df$year <- 2025
    
    df$name <- gsub(" ", "", df$name)
    
  df.spectra <- csv.spectra.AV %>%
    group_by(id) %>%
    summarise(red = weighted.mean(raw.value, srf$S2A_SR_AV_B4), nir = weighted.mean(raw.value, srf$S2A_SR_AV_B8), ndvi = (nir - red)/(nir + red)) 
    
  output <- df.spectra %>%
    inner_join(df, by = join_by(id == name)) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    select(-red, -nir, -c(picture, seagrassA, seagrassB, `sorting lab`, seagrass.cover.Laurant, `spectral camera`, spectra, type, habitat)) %>%
    rename(NDVI = ndvi)

  output
})
```

```{r biomassPreprocessingCB}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.CB <- local({
  df <- read_xlsx('data/biomass/CB/Cadiz_Biomass_C_cores.xlsx') %>%
    filter(Group == 'Leaves' | Group == 'Roots') %>%
    select(c(Group, Sample, `Dry mass (g)`)) %>%
    filter(!Sample %in% c('27','28','29','30')) %>% # planned sites that were not accomplished (10 of each type)
    pivot_wider(names_from = Group, values_from = `Dry mass (g)`) %>%
    mutate(site = case_when(# High cover
                          Sample < 10 ~ 'High',
                          Sample >= 10 & Sample < 18 ~ 'Mid',
                          Sample >= 18 ~ 'Bare'
                          ),
           A_per_m2 = convert_biomass_to_per_m2(as.numeric(Leaves)),
           B_per_m2 = convert_biomass_to_per_m2(as.numeric(Roots)),
           id = paste0("ST", sprintf("%02d", as.numeric(1:26)))
           ) %>%
    select(-c(Leaves, Roots, Sample)) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2')
    df$year <- 2025
    df$weight_per_m2[df$site == 'Bare'] <- 0 # bare areas do not have leaf or root biomass

  df.spectra <- csv.spectra.CB %>%
    group_by(id) %>%
    summarise(red = weighted.mean(raw.value, srf$S2A_SR_AV_B4), nir = weighted.mean(raw.value, srf$S2A_SR_AV_B8), ndvi = (nir - red)/(nir + red))  
    
  output <- df.spectra %>%
    group_by(id) %>%
    inner_join(df, by = join_by(id)) %>%
    select(-red, -nir) 

  output %>% 
    rename(NDVI = ndvi)
})
```

```{r biomassPreprocessingFinal}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass <- local({
  bb <- csv.biomass.BB
  bb$Location <- 'Bourgneuf_bay'
  # bb$site <- "natural"
  bb$site <- case_when(grepl("H", bb$id) ~ 'high',
                       grepl("L", bb$id) ~ 'low',
                       grepl('2018', bb$id) ~ '2018',
                       grepl('2019', bb$id) ~ '2019')
  
  av <- csv.biomass.AV
  av$Location <- 'Aveiro'
  
  cd <- csv.biomass.CB
  cd$Location <- 'Cadiz'
  
  rbind(bb, av, cd)
})

csv.biomass$site_location <- paste(csv.biomass$site, csv.biomass$Location, sep = "_") %>%
  factor(levels = c("2018_Bourgneuf_bay", "2019_Bourgneuf_bay", "low_Bourgneuf_bay", "high_Bourgneuf_bay", "natural_Aveiro", "rewilded_Aveiro", "Bare_Cadiz", "Mid_Cadiz", "High_Cadiz"), labels = c("Bourgneuf Bay (2018)", "Bourgneuf Bay (2019)", "Bourgneuf Bay (sparse)", "Bourgneuf Bay (dense)", "Ria de Aveiro (natural)", "Ria de Aveiro (rewilded)", "Cadiz Bay (Bare)", "Cadiz Bay (Mid)", "Cadiz Bay (High)"))

## Post-processing
csv.biomass <- csv.biomass %>%
  filter(!id %in% c("NR43", "RG5", "RG12", "RG13", "NR52", "NR44") %>% replace_na(TRUE)) # remove anomalous data; 

## Remove duplicate data from the different indices
csv.biomass <- csv.biomass %>%
  distinct()
```


### Carbon content data

The dried leaf (aboveground biomass, AGB) and root (belowground biomass, BGB) biomass were pulverised prior to the [measurement of total carbon (TC) and inorganic carbon (IC) content in fly ash](https://www.shimadzu.com/an/apl/22347/index.html) using a [SSM-5000A Shimadzu Solid Sample Combusion Unit](https://www.shimadzu.com/an/products/total-organic-carbon-analysis/toc-analysis-system/ssm-5000a/index.html). To measure the TC content in a sample, about 70 g of powdered biomass is combusted at 900&deg;C for 8 minutes. For the IC content measurement, about 300 g of powdered biomass is mixed with 0.9 ml of phosphorous acid (HPO~3~) and combusted at 300&deg;C until no carbon dioxide is detected. The measurements are scaled to the total weight of the samples using Equation \@ref(eq:carbon). The [difference between the TC and IC is the total organic carbon content (TOC)](https://www.shimadzu.com/an/apl/24168/index.html).

<!-- check formula -->
\begin{equation}
Carbon\,concentration\,(g\,per\,C/kg) = \frac{Measured\,Carbon\,(mg) \times Dry\,Weight\,(\%) \times 1000}{Wet\,weight\,of\,essay\,(mg)} (\#eq:carbon)
\end{equation}


### General relationship between reflectance and biomass

As the relationship established using the [Normalised Difference Vegetation Index](https://en.wikipedia.org/wiki/Normalized_difference_vegetation_index) derived from hyperspectral data will be applied onto the Sentinel-2 (S2) image, the hyperspectral data needs to be translated to be equivalent to S2. This was explored and optimised in two ways:

- Using the centre [@zoffoli2020] vs the [peak](https://sigoiry.github.io/Fieldtrip_Rewrite_Sept2025/) wavelength of S2A bands, based on the [S2 Spectral Response Function (SRF) Document](https://sentiwiki.copernicus.eu/web/document-library#DocumentLibrary-SENTINEL-2Documents), and using the S2 SRF.
- Testing three different vegetative indices, based on the comparison by @bargain2013
  - Normalised Difference Vegetative Index (NDVI)
  - Atmospherically-Resistant Vegetative Index (ARVI)
  - Modified Normalised Difference Index equivalent to the B5 and B6 bands of S2 (mND)

The values and indices derived from the three different methods of designating bands are similar. After deliberation, the SRF function for the S2A satellite was used. 

The indices performed similarly. NDVI is the only index here that uses only the 10m spatial resolution bands, unlike the ARVI and mND which rely on the 20m spatial resolution coastal band. Therefore, NDVI was selected as the optimal index.


### Accuracy Assessment

The adjusted R^2^, root mean square error (RMSE), and relative/normalised RMSE were used to assess the performance of the linear models. The adjusted R^2^ was obtained from the `stat_regline_equation()` function in the `ggpubr` package [@ggpubr]. RMSE was obtained from the `lm()` function in the `stats` package [@Rversion]. The relative RMSE was obtained by dividing the RMSE with the mean of the independent variable.

For the non-linear model, residual standard error (RSE) was used to inspect the fit of the exponential curve. the RSE was obtained from the `nls()` function in the `stats` package [@Rversion].

## Result

### Biomass

The matchup between the leaf and root biomass seemed linear in relationship up to about 100 g/m^2^, before a saturation was reached. There is a total of `r csv.biomass %>% filter(site_location == 'Cadiz Bay (High)') %>% filter(biomass == 'A_per_m2') %>% nrow()` saturated stations, all from Cadiz (Figure \@ref(fig:biomass)).

```{r biomass, fig.cap = paste("Scatterplot of leaf biomass (AGB) to root biomass (BGB) of _Zostera noltii_ across all sites. While this matchup seems to be linear for much of the dataset, there are ", csv.biomass %>% filter(site_location == 'Cadiz Bay (High)') %>% filter(biomass == 'A_per_m2') %>% nrow(), " stations in Cadiz with very high AGB and BGB that suggests a saturation. Thus, an exponential curve was fitted. For the linear model, the saturated points were removed to estimate the linear relationship up to saturation. Owing to some stations having no seagrass AGB and BGB, ", csv.biomass %>% filter(biomass == "B_per_m2" & weight_per_m2 == 0) %>% nrow()," data points are overlapping at the plot origin (0,0).", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.leaf.root <- csv.biomass %>%
  pivot_wider(names_from = biomass, values_from = weight_per_m2) 

## Compare AGB to BGB
xmax <- 810
ymax <- 170
plt.biomass.norm.lm <- ggplot(csv.biomass.leaf.root %>% filter(site_location != 'Cadiz Bay (High)'), aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: All except the exceptionally high biomass stations')
    
## NLS
# using LM to estimate starting parameters for NLS to run
init.c <- max(csv.biomass.leaf.root$B_per_m2, na.rm = TRUE) 
csv.biomass.0 <- lm(init.c - B_per_m2 ~ A_per_m2, data = csv.biomass.leaf.root)
csv.biomass.nls <- nls(B_per_m2 ~ a * (1 - exp(b*A_per_m2)), data = csv.biomass.leaf.root, start = list(a = csv.biomass.0$coefficients[1], b = csv.biomass.0$coefficients[2]), lower = c(a = 0, b = 0))
csv.biomass.nls.equation <- substitute(italic(y) == a * (1 - e^(b*x) ) * "," * ~italic("Residual SE") ~ "=" ~ rse, 
                                       list(x = "x", y = "y",
                                            a = summary(csv.biomass.nls)$coefficients[1, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            b = summary(csv.biomass.nls)$coefficients[2, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            rse =  summary(csv.biomass.nls)$sigma %>% round(2) %>% format(nsmall = 2)
                                            )) %>% 
  as.expression() %>%
  as.character()

plt.biomass.all.nls <- ggplot(csv.biomass.leaf.root, aes(x = A_per_m2, y = B_per_m2)) +
  geom_line(data = data.frame(x = csv.biomass.leaf.root$A_per_m2, y = predict(csv.biomass.nls, newdata = csv.biomass.leaf.root)), aes(x = x, y = y), colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  annotate("text", x = 1, y = Inf*0.9, hjust = 0, vjust = 1.2, label = csv.biomass.nls.equation, parse = TRUE, colour = 'black') +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Exponential model: All data')

girafe(ggobj = (plt.biomass.all.nls + plt.biomass.norm.lm) + plot_layout(guides = 'collect', axes = 'collect') + plot_annotation(title = "AGB vs BGB"),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 7, width_svg = 14
)
```

Thus, the saturated points were excluded from the dataset to obtain a linear relationship (Figure \@ref(fig:biomassFinal), Adj. R^2^ = 0.59, RMSE = 19.5). 

```{r biomassFinal, fig.cap = paste("A zoomed-in look at the scatterplot of leaf biomass (AGB) to root biomass (BGB) of _Zostera noltii_ across all sites, excluding the very high biomass (saturated) data points in Cadiz. The RMSE for this model is in the table below.", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

plt.biomass.norm.lm.rescaled <- ggplot(csv.biomass.leaf.root %>% filter(site_location != 'Cadiz Bay (High)'), aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, 150)) + ylim(c(0, 110)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: All except the exceptionally high biomass stations (rescaled)')

girafe(ggobj = plt.biomass.norm.lm.rescaled + plot_annotation(title = "AGB vs BGB"),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 5, width_svg = 10
)
```

```{r biomassFinalRMSE}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass %>%
  ungroup() %>%
  filter(site_location != 'Cadiz Bay (High)') %>%
  pivot_wider(names_from = biomass, values_from = weight_per_m2) %>%
  distinct() %>%
  drop_na(c(A_per_m2, B_per_m2)) %>%
  summarise(`Relative RMSE` = summary(lm(B_per_m2 ~ A_per_m2))$residuals^2 %>% mean() %>% sqrt() %>% "/"(mean(B_per_m2)) %>% "*"(100),
            RMSE = summary(lm(B_per_m2 ~ A_per_m2))$residuals^2 %>% mean() %>% sqrt())
```


### Hyperspectral data

For more about the hyperspectral information, please see [here](https://leechengfa.github.io/seagrass_reflectance_2025/).


### Carbon data

Work in progress. This section will be updated whenever the work is completed.



### General relationship between the reflectance and biomass

Similar to the matchup between AGB and BGB, there is a saturation of the NDVI at higher AGB (> 200 g/m^2^) [@bargain2013]. 

```{r reflectance, fig.cap = paste("Scatterplot of spectroradiometer-derived NDVI to seagrass leaf biomass. An exponential curve was fitted to the dataset. By removing the saturated points (AGB > 200 g/m^2^), a linear relationship can be obtained up to the estimated point of saturation.", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.reflectance <- csv.biomass %>%
  filter(biomass == 'A_per_m2') %>%
  distinct()

xmax <- 810
ymax <- 1

plt.raw_biomass_index.norm.lm <- ggplot(csv.biomass.reflectance %>% filter(weight_per_m2 < 200), aes(x = weight_per_m2, y = NDVI)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle(bquote('Linear model: AGB < 200'~g/m^2))
    
## NLS
# using LM to estimate starting parameters for NLS to run
init.d <- max(csv.biomass.reflectance$NDVI, na.rm = TRUE) 
csv.bioreflect.0 <- lm(init.d - NDVI ~ weight_per_m2, data = csv.biomass.reflectance)
csv.bioreflect.nls <- nls(NDVI ~ a * (1 - exp(b*weight_per_m2)), data = csv.biomass.reflectance, start = list(a = csv.bioreflect.0$coefficients[1], b = csv.bioreflect.0$coefficients[2]), lower = c(a = 0, b = 0))
csv.bioreflect.nls.equation <- substitute(italic(y) == a * (1 - e^(b*x) ) * "," * ~italic("Residual SE") ~ "=" ~ rse, 
                                       list(x = "x", y = "y",
                                            a = summary(csv.bioreflect.nls)$coefficients[1, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            b = summary(csv.bioreflect.nls)$coefficients[2, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            rse =  summary(csv.bioreflect.nls)$sigma %>% round(2) %>% format(nsmall = 2)
                                            )) %>% 
  as.expression() %>%
  as.character()

plt.raw_biomass_index.all.nls <- ggplot(csv.biomass.reflectance, aes(x = weight_per_m2, y = NDVI)) +
  geom_line(data = data.frame(x = csv.biomass.reflectance$weight_per_m2, y = predict(csv.bioreflect.nls, newdata = csv.biomass.reflectance)), aes(x = x, y = y), colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  annotate("text", x = 1, y = Inf*0.9, hjust = 0, vjust = 1.2, label = csv.bioreflect.nls.equation, parse = TRUE, colour = 'black') +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Exponential model: All data')

girafe(ggobj = (plt.raw_biomass_index.all.nls + plt.raw_biomass_index.norm.lm) + plot_layout(guides = 'collect', axes = 'collect') + plot_annotation(title = 'Spectroradiometer NDVI to leaf biomass'),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 7, width_svg = 11
)
```

A linear relationship was established between the spectroradiometer-derived NDVI and the seagrass leaf biomass (Figure \@ref(fig:reflectanceFinal), Adj. R^2^ = 0.63, RMSE = 0.120) up to 200 g/m^2^ in AGB. 

```{r reflectanceFinal, fig.cap = paste("A zoomed-in look at the scatterplot of spectroradiometer-derived NDVI to seagrass leaf biomass. The RMSE for this model is in the table below.", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.reflectance <- csv.biomass %>%
  filter(biomass == 'A_per_m2') %>%
  distinct()

plt.raw_biomass_index.norm.lm.rescaled <- ggplot(csv.biomass.reflectance %>% filter(weight_per_m2 < 200), aes(x = weight_per_m2, y = NDVI)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location', drop = FALSE) +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, 150)) + ylim(c(0, 1)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle(bquote('Linear model: AGB < 200'~g/m^2))
    
girafe(ggobj = plt.raw_biomass_index.norm.lm.rescaled + plot_annotation(title = 'Spectroradiometer NDVI to leaf biomass'),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 7, width_svg = 11
)
```

```{r reflectanceFinalRMSE}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass %>%
  ungroup() %>%
  filter(biomass == 'A_per_m2') %>%
  filter(weight_per_m2 < 200) %>%
  distinct() %>%
  drop_na(c(weight_per_m2, NDVI)) %>%
  summarise(`Relative RMSE` = summary(lm(NDVI ~ weight_per_m2))$residuals^2 %>% mean() %>% sqrt() %>% "/"(mean(NDVI)) %>% "*"(100),
            RMSE = summary(lm(NDVI ~ weight_per_m2))$residuals^2 %>% mean() %>% sqrt())
```

## References