---
title: Is there a general relationship between the seagrass _Zostera noltii_ biomass and the field
  spectroradiometer readings?
author: "Lee, Chengfa Benjamin"
date: "18 December 2025 (last updated on `r format(Sys.Date(), '%d %B %Y')`)"
output: bookdown::html_document2 # html_document (for publishing), bookdown::html_document2 (for publishing with dynamic referencing), html_notebook (for previewing)
bibliography: references.bib 
editor_options:
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(magrittr)
library(readxl)
library(zoo) 
library(ggplot2)
library(ggiraph)
library(dplyr)
library(tidyr)
library(stringr)
library(ggpubr)
library(ggtext)
library(patchwork)
```

This study aims to establish a relationship between the amount of biomass and reflectance of the _Zostera noltii_ seagrass. 

## Method

This data exploration uses the hyperspectral and biomass data collected during these field campaigns in Aveiro ([July 2025](https://sigoiry.github.io/Spectra_Aveiro_2025/)), Bourgneuf Bay ([2018, 2019](https://coastobs.eu/success-story/french-atlantic-coast), and [September 2025](https://sigoiry.github.io/Fieldtrip_Rewrite_Sept2025/)), and Cadiz Bay ([June 2025](https://rewriteproject.eu/)). 

### Hyperspectral data

For each sample, the hyperspectral signature was smoothed with a moving window of size 11 using the `rollapply()` function from the `zoo` package [@zoo2005]. While this means that the first and last five wavelengths would be smoothed with a partial moving window of less than 11 values, given the spectroradiometer's full wavelength range of 325nm to 1075nm and the final working range of 400nm to 900nm, these affected wavelengths will be filtered out and thus irrelevant here.

Post-smoothing, only the wavelengths from 400nm to 900nm are retained and a min-max standardisation was performed.

```{r hsPreprocessingBB}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.spectra.BB <- local({
  reflectance.2018 <- read.table('data/spectra/BB/2018_09_Bourgneuf_reflectance-cores_raw.txt', header = TRUE) %>%
    pivot_longer(cols = names(.)[-1], names_to = "id", values_to = 'raw.value') %>%
    rename(Wavelength = wavelength)
  
    reflectance.2018$id <- reflectance.2018$id %>%
      lapply(FUN = function(string) {
        string %>%
          strsplit(split = "\\.") %>% 
          unlist() %>%
          "["(1) %>%
          strsplit(split = "_") %>% 
          unlist() %>%
          "["(-length(.)) %>%
          paste(collapse = "_")
      }) %>%
      unlist()
    reflectance.2018$id <- paste(reflectance.2018$id, '2018', sep = ".")
    reflectance.2018$year <- 2018
    
  reflectance.2018 <- reflectance.2018 %>%
    group_by(id) %>%
    mutate(raw.value = rollapply(raw.value, 11, mean, partial = TRUE, align = 'center'), .keep = 'all') %>% 
    ungroup() %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    group_by(id) %>%
    mutate(std.value = (raw.value - min(raw.value))/(max(raw.value) - min(raw.value)), .keep = "all") %>%
    ungroup()
  
  reflectance.2025 <- read.csv('data/spectra/BB/20250911_Bourgneuf_H_stations_reflectance.txt', sep = '\t') %>%
    pivot_longer(cols = names(.) %>% grep('Wavelength', ., invert = TRUE, value = TRUE), names_to = "ASD", values_to = 'value') 
    
    reflectance.2025$id <- lapply(reflectance.2025$ASD, FUN = function(string) {
      str_split(string, '\\.') %>%
        unlist() %>%
        "["(1)
      }) %>% 
      unlist()
  
  reflectance.2025 <- reflectance.2025 %>%
    select(-ASD) %>%
    group_by(Wavelength,id) %>%
    summarise(raw.value = mean(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = rollapply(raw.value, 11, mean, partial = TRUE, align = 'center'), .keep = 'all') %>% 
    ungroup() %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    group_by(id) %>%
    mutate(std.value = (raw.value - min(raw.value))/(max(raw.value) - min(raw.value)), .keep = "all") %>%
    ungroup()
  
  reflectance.2025$id <- reflectance.2025$id %>%
    gsub('H', "H_", .)  
  reflectance.2025$year <- 2025
  
  reflectance.2018 %>% 
    rbind(reflectance.2025) %>%
    pivot_longer(cols = c(raw.value, std.value), names_to = 'processing', values_to = 'value')
})
```


```{r hsPreprocessingAV}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.spectra.AV <- local({
  df <- list.files('data/spectra/AV', full.names = TRUE) %>%
    grep('\\.txt', ., value = TRUE) %>%
    lapply(FUN = function(f) {
      g <- read.csv(f, sep = '\t')
      h <- names(g)[2] %>% 
        gsub('\\.asd', "", .)
        names(g)[2] <- "reading"
        g$Name <- h
      
      g
    }) %>%
    do.call(rbind, .)
  
  df$dataset <- df$Name %>%
    lapply(FUN = function(name) {
      out <- str_split(name, '00') %>% 
        unlist() %>%
        head(1) %>%
        unlist()
    }) %>% 
    unlist()
  df$rep <- df$Name %>%
    lapply(FUN = function(name) {
      out <- str_split(name, '00') %>% 
        unlist() %>%
        tail(1) %>%
        unlist()
    }) %>%
    unlist() %>%
    as.numeric() %>%
    "+"(1) 
  
  ## add in station IDs from metadata file
  df <- df %>%
    mutate(id = case_when(
    # NG group
    dataset == 'AVEI2' & rep >= 341 & rep <= 360 ~ "NG12",
    dataset == 'AVEI2' & rep >= 401 & rep <= 420 ~ "NG15",
    dataset == 'AVEI2' & rep >= 381 & rep <= 400 ~ "NG16",
    dataset == 'AVEI2' & rep >= 301 & rep <= 320 ~ "NG18",
    dataset == 'AVEI2' & rep >= 281 & rep <= 300 ~ "NG25",
    dataset == 'AVEI2' & rep >= 261 & rep <= 280 ~ "NG30",
    dataset == 'AVEI2' & rep >= 221 & rep <= 240 ~ "NG31",
    dataset == 'AVEI2' & rep >= 201 & rep <= 220 ~ "NG32",
    dataset == 'AVEI2' & rep >= 361 & rep <= 380 ~ "NR43", 
    dataset == 'AVEI2' & rep >= 321 & rep <= 340 ~ "NR44", 
    dataset == 'AVEI2' & rep >= 241 & rep <= 260 ~ "NR52", 

    # HG 50 group (spectra for heat shock paper -- Simon // not used for carbon/biomass)
    dataset == 'SHSI' & rep >= 301 & rep <= 320 ~ "HG_50_1",
    dataset == 'SHSI' & rep >= 321 & rep <= 340 ~ "HG_50_2",
    dataset == 'SHSI' & rep >= 341 & rep <= 360 ~ "HG_50_3",
    dataset == 'SHSI' & rep >= 361 & rep <= 380 ~ "HG_50_4",
    dataset == 'SHSI' & rep >= 381 & rep <= 400 ~ "HG_50_5",

    # HG SHSI group
    dataset == 'SHSI' & rep >= 1 & rep <= 20 ~ "HG_SHSI_1",
    dataset == 'SHSI' & rep >= 21 & rep <= 40 ~ "HG_SHSI_2",
    dataset == 'SHSI' & rep >= 41 & rep <= 60 ~ "HG_SHSI_3",
    dataset == 'SHSI' & rep >= 61 & rep <= 80 ~ "HG_SHSI_4",
    dataset == 'SHSI' & rep >= 81 & rep <= 100 ~ "HG_SHSI_5",
    dataset == 'SHSI' & rep >= 101 & rep <= 120 ~ "HG_SHSI_6",
    dataset == 'SHSI' & rep >= 121 & rep <= 140 ~ "HG_SHSI_7",
    dataset == 'SHSI' & rep >= 141 & rep <= 160 ~ "HG_SHSI_8",
    dataset == 'SHSI' & rep >= 161 & rep <= 180 ~ "HG_SHSI_9",
    dataset == 'SHSI' & rep >= 181 & rep <= 200 ~ "HG_SHSI_10",
    dataset == 'SHSI' & rep >= 201 & rep <= 220 ~ "HG_SHSI_11",
    dataset == 'SHSI' & rep >= 221 & rep <= 240 ~ "HG_SHSI_12",
    dataset == 'SHSI' & rep >= 241 & rep <= 260 ~ "HG_SHSI_13",
    dataset == 'SHSI' & rep >= 261 & rep <= 280 ~ "HG_SHSI_14",
    dataset == 'SHSI' & rep >= 281 & rep <= 300 ~ "HG_SHSI_15",
    
    # HG AVEI2 group
    dataset == 'AVEI2' & rep >= 1 & rep <= 20 ~ "HG_AVEI2_1",
    dataset == 'AVEI2' & rep >= 21 & rep <= 40 ~ "HG_AVEI2_2",
    dataset == 'AVEI2' & rep >= 41 & rep <= 60 ~ "HG_AVEI2_3",
    dataset == 'AVEI2' & rep >= 61 & rep <= 80 ~ "HG_AVEI2_4",
    dataset == 'AVEI2' & rep >= 81 & rep <= 100 ~ "HG_AVEI2_5",
    dataset == 'AVEI2' & rep >= 101 & rep <= 120 ~ "HG_AVEI2_6",
    dataset == 'AVEI2' & rep >= 121 & rep <= 140 ~ "HG_AVEI2_7",
    dataset == 'AVEI2' & rep >= 141 & rep <= 160 ~ "HG_AVEI2_8",
    dataset == 'AVEI2' & rep >= 161 & rep <= 180 ~ "HG_AVEI2_9",
    dataset == 'AVEI2' & rep >= 181 & rep <= 200 ~ "HG_AVEI2_10",
    dataset == 'AVEI2' & rep >= 201 & rep <= 220 ~ "HG_AVEI2_11",
    
    # RR/RG group
    dataset == 'AVE' & rep >= 1 & rep <= 20 ~ "RG12",
    dataset == 'AVE' & rep >= 21 & rep <= 40 ~ "RR36",
    dataset == 'AVE' & rep >= 41 & rep <= 60 ~ "RG5", 
    dataset == 'A' & rep >= 1 & rep <= 20 ~ "RG13",
    dataset == 'A' & rep >= 21 & rep <= 40 ~ "RG6", 
    dataset == 'A' & rep >= 41 & rep <= 60 ~ "RG7", 
    dataset == 'A' & rep >= 61 & rep <= 80 ~ "RG8", 
    dataset == 'A' & rep >= 101 & rep <= 120 ~ "RR46",
    dataset == 'A' & rep >= 121 & rep <= 140 ~ "RG16",
    dataset == 'A' & rep >= 141 & rep <= 160 ~ "RG23",
    dataset == 'A' & rep >= 161 & rep <= 180 ~ "RR45",
    dataset == 'A' & rep >= 181 & rep <= 200 ~ "RG22",
    
    # RW_TR group
    dataset == 'A' & rep >= 201 & rep <= 220 ~ "RW_TR_1",
    dataset == 'A' & rep >= 221 & rep <= 240 ~ "RW_TR_2",
    dataset == 'A' & rep >= 241 & rep <= 260 ~ "RW_TR_3",
    dataset == 'A' & rep >= 261 & rep <= 280 ~ "RW_TR_4",
    dataset == 'A' & rep >= 281 & rep <= 300 ~ "RW_TR_5",
    dataset == 'A' & rep >= 301 & rep <= 320 ~ "RW_TR_6",
    dataset == 'A' & rep >= 321 & rep <= 340 ~ "RW_TR_7",
    dataset == 'A' & rep >= 341 & rep <= 360 ~ "RW_TR_8",
    dataset == 'A' & rep >= 361 & rep <= 380 ~ "RW_TR_9",
    dataset == 'A' & rep >= 381 & rep <= 400 ~ "RW_TR_10"
  ))
  
  df <- df %>%
    # average the readings 
    filter(!is.na(id)) %>%
    group_by(id, Wavelength) %>%
    select(-c(rep, Name)) %>%
    summarise(raw.value = mean(reading), dataset = unique(dataset)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = rollapply(raw.value, 11, mean, partial = TRUE, align = 'center'), .keep = 'all') %>%
    ungroup() %>%
    filter(Wavelength >= 400 & Wavelength <= 900) %>%
    group_by(id) %>%
    mutate(std.value = (raw.value - min(raw.value))/(max(raw.value) - min(raw.value)), .keep = 'all') %>%
    ungroup() %>%
    pivot_longer(cols = c(raw.value, std.value), names_to = 'processing', values_to = 'value')
  
  meta <- list.files("data/spectra/AV", full.names = TRUE) %>%
    grep('Metadata_spectra', ., value = TRUE) %>%
    read_xlsx() %>%
    select(-c(Spectra)) 
  
  df <- df %>%
    left_join(meta, by = join_by(id == ID))
})
```

```{r hsPreprocessingCB}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.spectra.CB <- local({
  reflectance <- read.csv('data/spectra/CB/20250610_rw_cadiz_seagrass.txt', sep = '\t') %>%
    pivot_longer(cols = names(.) %>% grep('Wavelength', ., invert = TRUE, value = TRUE), names_to = "ASD", values_to = 'value') %>%
    mutate(group = unlist(substr(ASD, 1, 4)), rep = as.numeric(unlist(substr(ASD, 7, 10))))
  
  readMe <- read.csv('data/spectra/CB/20250610_rw_cadiz_seagrass_readme.txt', sep = ',') %>%
    mutate(group = substr(Spectrum_id_start, 1, 4), 
           id_start = as.numeric(unlist(substr(Spectrum_id_start, 7, 10))), id_end = as.numeric(unlist(substr(Spectrum_id_end, 7, 10)))
           )
    
    reflectance$id <- reflectance %>% 
      apply(MARGIN = 1, FUN = function(row) {
        output <- readMe$Station[readMe$group == row['group'] & readMe$id_start <= as.numeric(row['rep']) & readMe$id_end >= as.numeric(row['rep'])]
        ifelse(is.null(output), NA, output)
      }) %>% 
      unlist()
  
  reflectance <- reflectance %>%
    select(-ASD) %>%
    filter(!is.na(id)) %>%
    group_by(Wavelength,id) %>%
    summarise(raw.value = mean(value, na.rm = TRUE)) %>%
    ungroup() %>%
    group_by(id) %>%
    mutate(raw.value = rollapply(raw.value, 11, mean, partial = TRUE, align = 'center'), .keep = 'all') %>% 
    ungroup() %>%
    filter(Wavelength >= 400, Wavelength <= 900) %>%
    group_by(id) %>%
    mutate(std.value = (raw.value - min(raw.value))/(max(raw.value) - min(raw.value)), .keep = "all") %>%
    ungroup()
  
  reflectance$year <- 2025
  
  reflectance %>% 
    pivot_longer(cols = c(raw.value, std.value), names_to = 'processing', values_to = 'value')
})
```

### Biomass data

Full methods can be found [here](https://leechengfa.github.io/seagrass_carbon_2025/).

```{r biomassPreprocessingBB}
#| echo: false
#| error: false
#| message: false
#| warning: false

convert_biomass_to_per_m2 <- function(biomass) {
  core_area <- pi*((15/2)^2)
  biomass_per_m2 <- biomass*(100^2/core_area)
  
  return(biomass_per_m2)
}

csv.biomass.BB <- local({
  biomass.2018 <- read.csv('data/biomass/BB/2018_09_Bourgneuf_cores_biomass_percent_cover.csv', sep = '\t') %>%
    rename(id = Sample_ID, A_per_m2 = `Leaf_dry_weight_g`) %>%
    select(-Percent_cover)
    biomass.2018$A_per_m2 <- convert_biomass_to_per_m2(biomass.2018$A_per_m2)
    biomass.2018$B_per_m2 <- NA
    biomass.2018$year <- 2018
    biomass.2018$id <- paste(biomass.2018$id, '2018', sep = ".")
    
  spectra.2018 <- csv.spectra.BB %>%
    filter(year == 2018) %>%
    group_by(id) %>%
    filter(processing == "raw.value") %>% 
    select(-processing) %>%
    summarise(
      coastal.445.hs = value[Wavelength == 445], blue.520.hs = value[Wavelength == 520], red.670.hs = value[Wavelength == 670], 
      nir.701.hs = value[Wavelength == 701], nir.743.hs = value[Wavelength == 743], nir.800.hs = value[Wavelength == 800], 
      ndvi.s2peak.hs = (nir.800.hs - red.670.hs)/(nir.800.hs + red.670.hs), 
      mND_B5B6.s2peak.hs = (nir.743.hs - nir.701.hs)/(nir.743.hs + nir.701.hs - 2*coastal.445.hs),
      )  
  
  df.2018 <- biomass.2018 %>%
    left_join(spectra.2018, by = join_by(id == id)) %>%
    group_by(id)  %>%
    select(-contains('blue'),-contains('red'),-contains('coastal'),-contains('nir')) %>%
    pivot_longer(cols = contains('hs'), names_sep = '\\.', names_to = c('index','approach','.value')) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    select(-c(approach)) %>%
    rename(value = hs)
  
  biomass.2019 <- read_xlsx('data/biomass/BB/Data AGB BGB Bourgneuf 2019.xlsx') %>%
    rename(id = Sample_ID, A_per_m2 = `Leaf biomass (g DW m-2)`, B_per_m2 = `Rizhome biomass  (g DW m-2)`, value = NDVI
) %>%
    select(-`Seagrass coverage (%)`)
    biomass.2019$index <- 'ndvi'
    biomass.2019$year <- 2019
    
  df.2019 <- biomass.2019 %>%
    group_by(id) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2')
    df.2019$id <- paste(df.2019$id, '2019', sep = ".")
  
  biomass.2025 <- read_xlsx('data/biomass/BB/DryWeight_Cores.xlsx') %>%
    rename(id = Station)
  
    biomass.2025$A_per_m2 <- convert_biomass_to_per_m2(biomass.2025$A)
    biomass.2025$B_per_m2 <- convert_biomass_to_per_m2(biomass.2025$B)
    biomass.2025$year <- 2025
    
  spectra.2025 <- csv.spectra.BB %>%
    filter(year == 2025) %>%
    group_by(id) %>%
    filter(processing == "raw.value") %>% 
    select(-processing) %>%
    summarise(
      # S2 srf peak wavelength
      coastal.445.hs = value[Wavelength == 445], blue.520.hs = value[Wavelength == 520], red.670.hs = value[Wavelength == 670], 
      nir.701.hs = value[Wavelength == 701], nir.743.hs = value[Wavelength == 743], nir.800.hs = value[Wavelength == 800], 
      ndvi.s2peak.hs = (nir.800.hs - red.670.hs)/(nir.800.hs + red.670.hs), 
      # arvi.s2peak.hs = (nir.800.hs - (red.670.hs - (blue.520.hs - red.670.hs)))/(nir.800.hs + (red.670.hs - (blue.520.hs - red.670.hs))), 
      mND_B5B6.s2peak.hs = (nir.743.hs - nir.701.hs)/(nir.743.hs + nir.701.hs - 2*coastal.445.hs),
      # # centre wavelength of band range (Zoffoli et al. ,2020)
      # coastal.434.hs = value[Wavelength == 434], blue.495.hs = value[Wavelength == 495], red.665.hs = value[Wavelength == 665], 
      # nir.705.hs = value[Wavelength == 705], nir.740.hs = value[Wavelength == 740], nir.842.hs = value[Wavelength == 842], 
      # ndvi.s2mid.hs = (nir.842.hs - red.665.hs)/(nir.842.hs + red.665.hs), 
      # arvi.s2mid.hs = (nir.842.hs - (red.665.hs - (blue.495.hs - red.665.hs)))/(nir.842.hs + (red.665.hs - (blue.495.hs - red.665.hs))), 
      # mND_B5B6.s2mid.hs = (nir.740.hs - nir.705.hs)/(nir.740.hs + nir.705.hs - 2*coastal.434.hs)
      )  
  
  df.2025 <- biomass.2025 %>%
    left_join(spectra.2025, by = join_by(id == id)) %>%
    group_by(id)  %>%
    select(-contains('blue'),-contains('red'),-contains('coastal'),-contains('nir'), -A, -B) %>%
    pivot_longer(cols = contains('hs'), names_sep = '\\.', names_to = c('index','approach','.value')) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    # filter(approach == "s2peak") %>%
    # filter(index != 'arvi') %>%
    select(-c(approach)) %>%
    rename(value = hs)
  
  rbind(df.2018, df.2019, df.2025)
})
```

```{r biomassPreprocessingAV}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.AV <- local({
  df <- read_xlsx('data/biomass/AV/Aveiro_grid_info_final.xlsx', sheet = 'data') %>%
    select(-c(longitude, latitude, poskey.benthos, poskey.sediment, remark))
    df$`spectral camera`[df$`spectral camera` == 'no'] <- NA
    df$spectra <- !is.na(df$`spectral camera`)
    
    df$A_per_m2 <- convert_biomass_to_per_m2(df$seagrassA/1000)
    df$B_per_m2 <- convert_biomass_to_per_m2(df$seagrassB/1000)
    df$A_per_m2[df$B_per_m2 == 0] <- 0 # if there's no root biomass, the leaf biomass are not growing at this station
    df$year <- 2025
    
    df$name <- gsub(" ", "", df$name)
    
  df.spectra <- csv.spectra.AV %>%
    group_by(id) %>%
    filter(processing == "raw.value") %>% 
    select(-processing) %>%
    mutate(
      # S2 srf peak wavelength
      coastal.445.hs = value[Wavelength == 445], blue.520.hs = value[Wavelength == 520], red.670.hs = value[Wavelength == 670], 
      nir.701.hs = value[Wavelength == 701], nir.743.hs = value[Wavelength == 743], nir.800.hs = value[Wavelength == 800], 
      ndvi.s2peak.hs = (nir.800.hs - red.670.hs)/(nir.800.hs + red.670.hs), 
      arvi.s2peak.hs = (nir.800.hs - (red.670.hs - (blue.520.hs - red.670.hs)))/(nir.800.hs + (red.670.hs - (blue.520.hs - red.670.hs))), 
      mND_B5B6.s2peak.hs = (nir.743.hs - nir.701.hs)/(nir.743.hs + nir.701.hs - 2*coastal.445.hs),
      # centre wavelength of band range (Zoffoli et al. ,2020)
      coastal.434.hs = value[Wavelength == 434], blue.495.hs = value[Wavelength == 495], red.665.hs = value[Wavelength == 665], 
      nir.705.hs = value[Wavelength == 705], nir.740.hs = value[Wavelength == 740], nir.842.hs = value[Wavelength == 842], 
      ndvi.s2mid.hs = (nir.842.hs - red.665.hs)/(nir.842.hs + red.665.hs), 
      arvi.s2mid.hs = (nir.842.hs - (red.665.hs - (blue.495.hs - red.665.hs)))/(nir.842.hs + (red.665.hs - (blue.495.hs - red.665.hs))), 
      mND_B5B6.s2mid.hs = (nir.740.hs - nir.705.hs)/(nir.740.hs + nir.705.hs - 2*coastal.434.hs)
      ) %>%
    ungroup()
    
  output <- df.spectra %>%
    group_by(id) %>%
    summarise(dataset = unique(dataset), Group = unique(Group), Site = unique(Site), across(names(df.spectra %>% select(-c(id,Wavelength,dataset,Group,Site,value))), ~ mean(.x, na.rm = TRUE))) %>%
    inner_join(df, by = join_by(id == name)) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2') %>%
    select(-c(SPC_expected, SPC_TRUE, Benthos_code, Sediment_code,picture, seagrassA, seagrassB, `sorting lab`, seagrass.cover.Laurant, `spectral camera`, spectra)) %>%
    select(-contains('blue'),-contains('red'),-contains('coastal'),-contains('nir')) %>%
    pivot_longer(cols = contains('hs'), names_sep = '\\.', names_to = c('index','approach','.value'))

  output %>% 
    filter(approach == "s2peak") %>%
    filter(biomass != 'T_per_m2') %>%
    filter(index != 'arvi') %>%
    select(-c(approach, habitat, Group, dataset, type, Site)) %>%
    rename(value = hs) 
})
```

```{r biomassPreprocessingCB}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.CB <- local({
  df <- read_xlsx('data/biomass/CB/Cadiz_Biomass_C_cores.xlsx') %>%
    filter(Group == 'Leaves' | Group == 'Roots') %>%
    select(c(Group, Sample, `Dry mass (g)`)) %>%
    filter(!Sample %in% c('27','28','29','30')) %>% # planned sites that were not accomplished (10 of each type)
    pivot_wider(names_from = Group, values_from = `Dry mass (g)`) %>%
    mutate(site = case_when(# High cover
                          Sample < 10 ~ 'High',
                          Sample >= 10 & Sample < 18 ~ 'Mid',
                          Sample >= 18 ~ 'Bare'
                          ),
           A_per_m2 = convert_biomass_to_per_m2(as.numeric(Leaves)),
           B_per_m2 = convert_biomass_to_per_m2(as.numeric(Roots)),
           id = paste0("ST", sprintf("%02d", as.numeric(1:26)))
           ) %>%
    select(-c(Leaves, Roots, Sample)) %>%
    pivot_longer(cols = c(A_per_m2, B_per_m2), names_to = 'biomass', values_to = 'weight_per_m2')
    df$year <- 2025

  df.spectra <- csv.spectra.CB %>%
    group_by(id) %>%
    filter(processing == "raw.value") %>% 
    select(-processing) %>%
    mutate(
      # S2 srf peak wavelength
      coastal.445.hs = value[Wavelength == 445], blue.520.hs = value[Wavelength == 520], red.670.hs = value[Wavelength == 670], 
      nir.701.hs = value[Wavelength == 701], nir.743.hs = value[Wavelength == 743], nir.800.hs = value[Wavelength == 800], 
      ndvi.s2peak.hs = (nir.800.hs - red.670.hs)/(nir.800.hs + red.670.hs), 
      arvi.s2peak.hs = (nir.800.hs - (red.670.hs - (blue.520.hs - red.670.hs)))/(nir.800.hs + (red.670.hs - (blue.520.hs - red.670.hs))), 
      mND_B5B6.s2peak.hs = (nir.743.hs - nir.701.hs)/(nir.743.hs + nir.701.hs - 2*coastal.445.hs),
      # centre wavelength of band range (Zoffoli et al. ,2020)
      coastal.434.hs = value[Wavelength == 434], blue.495.hs = value[Wavelength == 495], red.665.hs = value[Wavelength == 665], 
      nir.705.hs = value[Wavelength == 705], nir.740.hs = value[Wavelength == 740], nir.842.hs = value[Wavelength == 842], 
      ndvi.s2mid.hs = (nir.842.hs - red.665.hs)/(nir.842.hs + red.665.hs), 
      arvi.s2mid.hs = (nir.842.hs - (red.665.hs - (blue.495.hs - red.665.hs)))/(nir.842.hs + (red.665.hs - (blue.495.hs - red.665.hs))), 
      mND_B5B6.s2mid.hs = (nir.740.hs - nir.705.hs)/(nir.740.hs + nir.705.hs - 2*coastal.434.hs)
      ) %>%
    ungroup()
    
  output <- df.spectra %>%
    group_by(id) %>%
    summarise(across(names(df.spectra %>% select(-c(id,Wavelength,value))), ~ mean(.x, na.rm = TRUE))) %>%
    inner_join(df, by = join_by(id)) %>%
    select(-contains('blue'),-contains('red'),-contains('coastal'),-contains('nir')) %>%
    pivot_longer(cols = contains('hs'), names_sep = '\\.', names_to = c('index','approach','.value'))

  output %>% 
    select(-c(approach)) %>%
    rename(value = hs) 
})
```

```{r biomassPreprocessingFinal}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass <- local({
  bb <- csv.biomass.BB
  bb$Location <- 'Bourgneuf_bay'
  # bb$site <- "natural"
  bb$site <- case_when(grepl("H", bb$id) ~ 'high',
                       grepl("L", bb$id) ~ 'low',
                       grepl('2018', bb$id) ~ '2018',
                       grepl('2019', bb$id) ~ '2019')
  
  av <- csv.biomass.AV
  av$Location <- 'Aveiro'
  
  cd <- csv.biomass.CB
  cd$Location <- 'Cadiz'
  
  rbind(bb, av, cd)
})

csv.biomass$site_location <- paste(csv.biomass$site, csv.biomass$Location, sep = "_") %>%
  factor(levels = c("2018_Bourgneuf_bay", "2019_Bourgneuf_bay", "low_Bourgneuf_bay", "high_Bourgneuf_bay", "natural_Aveiro", "rewilded_Aveiro", "Bare_Cadiz", "Mid_Cadiz", "High_Cadiz"), labels = c("Bourgneuf Bay (2018)", "Bourgneuf Bay (2019)", "Bourgneuf Bay (sparse)", "Bourgneuf Bay (dense)", "Ria de Aveiro (natural)", "Ria de Aveiro (rewilded)", "Cadiz Bay (Bare)", "Cadiz Bay (Mid)", "Cadiz Bay (High)"))

## Post-processing
csv.biomass <- csv.biomass %>%
  filter(!id %in% c("NR43", "RG5", "RG12", "RG13", "NR52", "NR44") %>% replace_na(TRUE)) # remove anomalous data; 

## Remove duplicate data from the different indices
csv.biomass <- csv.biomass %>%
  filter(index == 'ndvi') %>%
  select(-index) %>% 
  distinct()

# print(csv.biomass %>% select(id) %>% table()) ## for counting data. Each station should have 2 points in the frequency table (AGB & BGB).

```


### Carbon content data

The dried leaf (aboveground biomass, AGB) and root (belowground biomass, BGB) biomass were pulverised prior to the [measurement of total carbon (TC) and inorganic carbon (IC) content in fly ash](https://www.shimadzu.com/an/apl/22347/index.html) using a [SSM-5000A Shimadzu Solid Sample Combusion Unit](https://www.shimadzu.com/an/products/total-organic-carbon-analysis/toc-analysis-system/ssm-5000a/index.html). To measure the TC content in a sample, about 70 g of powdered biomass is combusted at 900&deg;C for 8 minutes. For the IC content measurement, about 300 g of powdered biomass is mixed with 0.9 ml of phosphorous acid (HPO~3~) and combusted at 300&deg;C until no carbon dioxide is detected. The measurements are scaled to the total weight of the samples using Equation \@ref(eq:carbon). The [difference between the TC and IC is the total organic carbon content (TOC)](https://www.shimadzu.com/an/apl/24168/index.html).

<!-- check formula -->
\begin{equation}
Carbon\,concentration\,(g\,per\,C/kg) = \frac{Measured\,Carbon\,(mg) \times Dry\,Weight\,(\%) \times 1000}{Wet\,weight\,of\,essay\,(mg)} (\#eq:carbon)
\end{equation}


### General relationship between reflectance and biomass

As the relationship established using the hyperspectral data will be applied onto the Sentinel-2 (S2) image, the hyperspectral data needs to be translated to be equivalent to S2. This was explored and optimised in two ways:

- Using the centre [@zoffoli2020] vs the [peak](https://sigoiry.github.io/Fieldtrip_Rewrite_Sept2025/) wavelength of S2A bands, based on the [S2 Spectral Response Function Document: COPE-GSEG-EOPG-TN-15-0007](https://sentiwiki.copernicus.eu/web/document-library#DocumentLibrary-SENTINEL-2Documents)
- Testing three different vegetative indices, based on the comparison by @bargain2013.
  - Normalised Difference Vegetative Index (NDVI)
  - Atmospherically-Resistant Vegetative Index (ARVI)
  - Modified Normalised Difference Index equivalent to the B5 and B6 bands of S2 (mND)

While there were no significant difference in predictive performance between the centre and peak band wavelength, the latter was slightly better. For brevity, the results henceforth are based on the peak wavelength of the S2 bands.

Similarly, the indices performed similarly. NDVI is the only index here that uses only the 10m spatial resolution bands, unlike the ARVI and mND which rely on the 20m spatial resolution coastal band. Therefore, NDVI was selected as the optimal index.


### Accuracy Assessment

The adjusted R^2^, root mean square error (RMSE), and relative/normalised RMSE were used to assess the performance of the linear models. The adjusted R^2^ was obtained from the `stat_regline_equation()` function in the `ggpubr` package [@ggpubr]. RMSE was obtained from the `lm()` function in the `stats` package [@Rversion]. The relative RMSE was obtained by dividing the RMSE with the mean of the independent variable.

## Result

### Biomass

While most of the dataset seemed linear in relationship, `r csv.biomass %>% filter(grepl('High', id)) %>% nrow()` stations in Cadiz stood out (Figure \@ref(fig:biomass).

```{r biomass, fig.cap = paste("Scatterplot of leaf biomass (AGB) to root biomass (BGB) of _Zostera noltii_ across all sites. While this matchup seems to be linear for much of the dataset, there are ", csv.biomass %>% filter(grepl('High', id)) %>% nrow(), " stations in Cadiz with very high AGB and BGB that suggests a saturation. As such, the dataset was split along this difference and modelled separately. Also, an exponential model was fitted on the full dataset to test if it could adequately describe the dataset. Owing to some stations having no seagrass AGB and BGB, ", csv.biomass %>% filter(biomass == "B_per_m2" & weight_per_m2 == 0) %>% nrow()," data points are overlapping at the plot origin (0,0).", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.leaf.root <- csv.biomass %>%
  pivot_wider(names_from = biomass, values_from = weight_per_m2) 

## Compare AGB to BGB
xmax <- 810
ymax <- 170
plt.biomass.all.lm <- ggplot(csv.biomass.leaf.root, aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location') +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: All data')
    
plt.biomass.norm.lm <- ggplot(csv.biomass.leaf.root %>% filter(!grepl('High_Cadiz', site)), aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: All except the exceptionally high biomass stations')
    
plt.biomass.veryhigh.lm <- ggplot(csv.biomass.leaf.root %>% filter(grepl('High_Cadiz', site)), aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: Only the exceptionally high biomass stations')

## NLS
# using LM to estimate starting parameters for NLS to run
init.c <- max(csv.biomass.leaf.root$B_per_m2, na.rm = TRUE) 
csv.biomass.0 <- lm(init.c - B_per_m2 ~ A_per_m2, data = csv.biomass.leaf.root)
csv.biomass.nls <- nls(B_per_m2 ~ a * (1 - exp(b*A_per_m2)), data = csv.biomass.leaf.root, start = list(a = csv.biomass.0$coefficients[1], b = csv.biomass.0$coefficients[2]), lower = c(a = 0, b = 0))
csv.biomass.nls.equation <- substitute(italic(y) == a * (1 - e^(b*x) ) * "," * ~italic("Residual SE") ~ "=" ~ rse, 
                                       list(x = "x", y = "y",
                                            a = summary(csv.biomass.nls)$coefficients[1, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            b = summary(csv.biomass.nls)$coefficients[2, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            rse =  summary(csv.biomass.nls)$sigma %>% round(2) %>% format(nsmall = 2)
                                            )) %>% 
  as.expression() %>%
  as.character()

plt.biomass.all.nls <- ggplot(csv.biomass.leaf.root, aes(x = A_per_m2, y = B_per_m2)) +
  geom_line(data = data.frame(x = csv.biomass.leaf.root$A_per_m2, y = predict(csv.biomass.nls, newdata = csv.biomass.leaf.root)), aes(x = x, y = y), colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location') +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  annotate("text", x = 1, y = Inf*0.9, hjust = 0, vjust = 1.2, label = csv.biomass.nls.equation, parse = TRUE, colour = 'black') +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Exponential model: All data')

girafe(ggobj = (plt.biomass.all.lm + plt.biomass.all.nls) / (plt.biomass.norm.lm + plt.biomass.veryhigh.lm) + plot_layout(guides = 'collect', axes = 'collect') + plot_annotation(title = "AGB vs BGB"),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 7, width_svg = 14
)
```

The aggregated dataset excluding the very high biomass stations suggests a linear relationship between the leaf and root biomass, with a large variation (Figure \@ref(fig:biomassFinal), Adj. R^2^ = 0.54, RMSE = 20.0). 

```{r biomassFinal, fig.cap = paste("Scatterplot of leaf biomass (AGB) to root biomass (BGB) of _Zostera noltii_ across all sites, excluding the very high biomass data points in Cadiz.", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

plt.biomass.norm.lm.rescaled <- ggplot(csv.biomass.leaf.root %>% filter(!grepl('High_Cadiz', site)), aes(x = A_per_m2, y = B_per_m2)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  ylab('Belowground Biomass (g/m<sup>2</sup>)') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, 150)) + ylim(c(0, 110)) +
  scale_colour_discrete('Location') +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Linear model: All except the exceptionally high biomass stations')

girafe(ggobj = plt.biomass.norm.lm.rescaled + plot_annotation(title = "AGB vs BGB"),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 5, width_svg = 10
)
```

```{r biomassFinalRMSE}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass %>%
  ungroup() %>%
  filter(!grepl('High_Cadiz', site_location)) %>%
  pivot_wider(names_from = biomass, values_from = weight_per_m2) %>%
  distinct() %>%
  drop_na(c(A_per_m2, B_per_m2)) %>%
  summarise(`Relative RMSE` = summary(lm(B_per_m2 ~ A_per_m2))$residuals^2 %>% mean() %>% sqrt() %>% "/"(mean(A_per_m2)) %>% "*"(100),
            RMSE = summary(lm(B_per_m2 ~ A_per_m2))$residuals^2 %>% mean() %>% sqrt())
```


### Hyperspectral data

```{r hyperspectral_preprocessing}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.hs <- local({
  csv.spectra.BB$Location <- 'Bourgneuf_bay'
  
  csv.spectra.AV$Location <- 'Aveiro'
  csv.spectra.AV2 <- csv.spectra.AV  %>% 
    select(-c(contains("hs"), contains("SPC"), Benthos_code, Sediment_code, Group, dataset, Site)) 
    csv.spectra.AV2$year <- 2025
    
  csv.spectra.CB$Location <- 'Cadiz'
  
  rbind(csv.spectra.BB, csv.spectra.AV2, csv.spectra.CB) 
})

csv.hs.long <- csv.hs %>%
  group_by(id) %>%
  left_join(csv.biomass %>% select(c(id, biomass, weight_per_m2, site_location)), by = join_by(id == id), relationship = 'many-to-many') %>% 
  distinct() %>%
  filter((biomass != 'B_per_m2') %>% replace_na(TRUE)) 
```

Generally, the greater the dry weight, the lower the reflectance (Figure \@ref(fig:hsRaw) and Figure \@ref(fig:hsStd)). In particular, the absorption at around 670nm (S2 red band) is greater with more seagrass AGB. 


```{r hsRaw, fig.cap = "Raw spectroradiometer reflectance measurements compared against the dry weight (g/m^2^). There were no hyperspectral readings taken in the sparse meadows in Bourgneuf Bay."}
#| echo: false
#| error: false
#| message: false
#| warning: false

plt.raw_biomass_comparison <- csv.hs.long %>%
  filter(!is.na(site_location)) %>%
  filter(processing == 'raw.value') %>%
  ggplot(aes(x = Wavelength, y = value, colour = weight_per_m2, group = id)) + 
  geom_line_interactive(aes(tooltip = paste0("AGB: ", round(weight_per_m2, 2), ' g/m<sup>2</sup> (id: ', id,')'), data_id = id)) +
  scale_colour_viridis_c(option = 'turbo') +
  xlab('Wavelength (nm)') + ylab('Reflectance') + labs(colour = 'Dry Weight (g/m<sup>2</sup>)') +
  facet_grid(. ~ site_location) +
  theme(legend.title = element_markdown()) +
  ggtitle('Raw values of spectroradiometer measurements to the leaf dry weight')

girafe(ggobj = plt.raw_biomass_comparison,
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 5, width_svg = 14
)
```

```{r hsStd, fig.cap = "Min-max standardised spectroradiometer reflectance measurements compared against the dry weight (g/m^2^). There were no hyperspectral readings taken in the sparse meadows in Bourgneuf Bay."}
#| echo: false
#| error: false
#| message: false
#| warning: false

plt.std_biomass_comparison <- csv.hs.long %>%
  drop_na() %>%
  filter(processing == 'std.value') %>%
  ggplot(aes(x = Wavelength, y = value, colour = weight_per_m2, group = id)) + 
  geom_line_interactive(aes(tooltip = paste0("AGB: ", round(weight_per_m2, 2), ' g/m<sup>2</sup> (id: ', id,')'), data_id = id)) +
  scale_colour_viridis_c(option = 'turbo') +
  xlab('Wavelength (nm)') + ylab('Reflectance') + labs(colour = 'Dry Weight (g/m<sup>2</sup>)') +
  facet_grid(. ~ site_location) +
  theme(legend.title = element_markdown()) +
  ggtitle('Min-max standardised values of spectroradiometer measurements to the leaf dry weight')

girafe(ggobj = plt.std_biomass_comparison,
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 5, width_svg = 14
)
```

### Carbon data

Work in progress. This section will be updated whenever the work is completed.



### General relationship between the reflectance and biomass

A linear relationship was established between the spectroradiometer-derived NDVI and the seagrass leaf biomass (Figure \@ref(fig:reflectance), Adj. R^2^ = 0.68, RMSE = 20.4). The points furthest from the established linear equation are RR36 (Aveiro rewilded), RR45 (Aveiro rewilded) and H_23 (Bourgneuf Bay dense meadow).

```{r reflectance, fig.cap = paste("Scatterplot of spectroradiometer-derived NDVI to seagrass leaf biomass.", sep = "")}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass.reflectance <- csv.biomass %>%
  filter(biomass == 'A_per_m2') %>%
  distinct()

xmax <- 810
ymax <- 1

plt.raw_biomass_index.all.lm <- csv.biomass.reflectance %>%
  ggplot(aes(x = weight_per_m2, y = value)) + 
    geom_smooth(method = 'lm', colour = 'black') +
    geom_point_interactive(aes(tooltip = paste0("Biomass: ", round(weight_per_m2, 2), ' g/m<sup>2</sup> (id: ', id,')'), data_id = id, colour = site_location)) +
    ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
    stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
    scale_colour_discrete('Location') +
    theme(axis.title.y = element_markdown()) +
    ggtitle('Linear model: All data')

plt.raw_biomass_index.norm.lm <- ggplot(csv.biomass.reflectance %>% filter(weight_per_m2 < 200), aes(x = weight_per_m2, y = value)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle(bquote('Linear model: AGB > 200'~g/m^2))
    
plt.raw_biomass_index.veryhigh.lm <- ggplot(csv.biomass.reflectance %>% filter(weight_per_m2 > 200), aes(x = weight_per_m2, y = value)) +
  geom_smooth(method = 'lm', colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location), show.legend = FALSE) +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  stat_regline_equation(label.x.npc = 'left', label.y.npc = 'top', decreasing = TRUE, aes(label =  paste(..eq.label.., ..adj.rr.label.., sep = "~~~~"))) +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle(bquote('Linear model: AGB > 200'~g/m^2))


## NLS
# using LM to estimate starting parameters for NLS to run
init.d <- max(csv.biomass.reflectance$value, na.rm = TRUE) 
csv.bioreflect.0 <- lm(init.d - value ~ weight_per_m2, data = csv.biomass.reflectance)
csv.bioreflect.nls <- nls(value ~ a * (1 - exp(b*weight_per_m2)), data = csv.biomass.reflectance, start = list(a = csv.bioreflect.0$coefficients[1], b = csv.bioreflect.0$coefficients[2]), lower = c(a = 0, b = 0))
csv.bioreflect.nls.equation <- substitute(italic(y) == a * (1 - e^(b*x) ) * "," * ~italic("Residual SE") ~ "=" ~ rse, 
                                       list(x = "x", y = "y",
                                            a = summary(csv.bioreflect.nls)$coefficients[1, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            b = summary(csv.bioreflect.nls)$coefficients[2, 'Estimate'] %>% unname() %>% round(2) %>% format(nsmall = 2),
                                            rse =  summary(csv.bioreflect.nls)$sigma %>% round(2) %>% format(nsmall = 2)
                                            )) %>% 
  as.expression() %>%
  as.character()

plt.raw_biomass_index.all.nls <- ggplot(csv.biomass.reflectance, aes(x = weight_per_m2, y = value)) +
  geom_line(data = data.frame(x = csv.biomass.reflectance$weight_per_m2, y = predict(csv.bioreflect.nls, newdata = csv.biomass.reflectance)), aes(x = x, y = y), colour = 'black') +
  geom_point_interactive(aes(tooltip = paste0('id: ', id), data_id = id, colour = site_location, shape = Location)) +
  scale_shape_discrete('Site', labels = c('Aveiro' = 'Ria de Aveiro', 'Bourgneuf_bay' = 'Bourgneuf Bay', 'Cadiz' = 'Cadiz Bay')) +
  scale_colour_discrete('Location') +
  ylab('NDVI') + xlab('Aboveground Biomass (g/m<sup>2</sup>)') +
  xlim(c(0, xmax)) + ylim(c(0, ymax)) +
  annotate("text", x = 1, y = Inf*0.9, hjust = 0, vjust = 1.2, label = csv.bioreflect.nls.equation, parse = TRUE, colour = 'black') +
  theme(axis.title.x = element_markdown(), axis.title.y = element_markdown()) +
  ggtitle('Exponential model: All data')

girafe(ggobj = (plt.raw_biomass_index.all.lm + plt.raw_biomass_index.all.nls) / (plt.raw_biomass_index.norm.lm + plt.raw_biomass_index.veryhigh.lm) + plot_layout(guides = 'collect', axes = 'collect') + plot_annotation(title = 'Spectroradiometer NDVI to leaf biomass'),
       options = list(
         opts_hover(css = "stroke-width:2px"),
         opts_hover_inv(css = "opacity:0.1"),
         opts_tooltip(opacity = 0.95, css = "padding:6px; font-size:12px")
       ),
       height_svg = 7, width_svg = 11
)
```

```{r reflectanceRMSE}
#| echo: false
#| error: false
#| message: false
#| warning: false

csv.biomass %>%
  ungroup() %>%
  filter(biomass == 'A_per_m2') %>%
  filter(weight_per_m2 > 200) %>%
  distinct() %>%
  drop_na(c(weight_per_m2, value)) %>%
  summarise(`Relative RMSE` = summary(lm(weight_per_m2 ~ value))$residuals^2 %>% mean() %>% sqrt() %>% "/"(mean(weight_per_m2)) %>% "*"(100),
            RMSE = summary(lm(weight_per_m2 ~ value))$residuals^2 %>% mean() %>% sqrt())
```

## References